---
title: "Tidy Tuesday 2022"
output: html_notebook
---

# 2022-01-18
```{r}
require(tidyverse)
# tuesdata <- tidytuesdayR::tt_load('2022-01-18')
# chocolate <- tuesdata$chocolate
# chocolate %>% write_csv('data/chocolate.csv')
chocolate <- read_csv('data/chocolate.csv',
                      col_types = cols(.default=col_character(),
                                       ref=col_number(),
                                       review_date=col_number(),
                                       rating=col_number()))
chocolate %>% head()
```


### Data Dictionary

# `chocolate.csv`

|variable                         |class     |description |
|:--------------------------------|:---------|:-----------|
|ref                              |integer   |Reference ID, The highest REF numbers were the last entries made. |
|company_manufacturer             |character | Manufacturer name |
|company_location                 |character | Manufacturer region |
|review_date                      |integer   | Review date (year) |
|country_of_bean_origin           |character | Country of origin |
|specific_bean_origin_or_bar_name |character | Specific bean or bar name|
|cocoa_percent                    |character | Cocoa percent (% chocolate) |
|ingredients                      |character | Ingredients, ("#" = represents the number of ingredients in the chocolate; B = Beans, S = Sugar, S* = Sweetener other than white cane or beet sugar, C = Cocoa Butter, V = Vanilla, L = Lecithin, Sa = Salt) |
|most_memorable_characteristics   |character | Most Memorable Characteristics column is a summary review of the most memorable characteristics of that bar. Terms generally relate to anything from texture, flavor, overall opinion, etc. separated by ','|
|rating                           |double    | rating between 1-5 |

```{r}
chocolate <- chocolate %>% 
  separate(ingredients, into=c('n_ingredients', 'ingredients'),
           sep='-') %>%
  mutate(n_ingredients = as.numeric(n_ingredients)) %>%
  mutate(ingredients = str_trim(ingredients)) %>%
  mutate(cocoa_val = str_sub(cocoa_percent, 1,
                                     nchar(cocoa_percent)-1)) %>%
  mutate_at(vars(cocoa_val), as.double)
```


```{r}
chocolate %>% 
  count(company_location) %>%
  ggplot(aes(n)) + geom_histogram(bins=30)
chocolate %>% count(company_location) %>% arrange(-n)
```
```{r}
common_producers <- chocolate %>% 
  count(company_location) %>% 
  filter(n >= 30) %>%
  select(company_location) %>%
  pull()
```


```{r}
chocolate %>% 
  count(country_of_bean_origin) %>%
  ggplot(aes(n)) + geom_histogram(bins=30)
chocolate %>% count(country_of_bean_origin) %>% arrange(-n)
```


```{r}
chocolate <- chocolate %>%
  mutate(region = case_when(
    company_location %in% c('Amsterdam',  'Czech Republic', 'Finland','Hungary',
                            'Iceland', 'Ireland', 'Lithuania', 'Netherlands', 
                            'Norway', 'Poland', 'Portugal',
                            'Scotland', 'Sweden', 'Wales') ~ 'Europe',
    company_location %in% c('Argentina', 'Bolivia', 'Brazil', 'Chile',
                            'Colombia', 'Costa Rica', 'Dominican Republic',
                            'El Salvador', 'Guatemala', 'Honduras',
                            'Nicaragua', 'Peru', 'Suriname') ~ 'South America',
    company_location %in% c('Fiji', 'New Zealand', 'Philippines', 'Vanuatu') ~ 'Oceania',
    company_location %in% c('Ghana', 'Madagascar', 'Malaysia', 'Sao Tome', 
                            'Sao Tome & Principe', 'South Africa') ~ 'Africa',
    company_location %in% c('Israel', "U.A.E.") ~ 'Middle East',
    company_location %in% c('Russia', 'Vietnam', 'India', 'Singapore',
                            'South Korea', 'Taiwan', 'Thailand') ~ 'Asia',
    company_location %in% c('Grenada', 'Puerto Rico', 'Martinique',
                            'St. Lucia', 'St.Vincent-Grenadines' ) ~ 'Caribbean',
    TRUE ~ company_location
  ))
```

```{r}
chocolate %>%
  select(cocoa_val, rating, n_ingredients, review_date) %>%
  GGally::ggpairs()
```
```{r}
chocolate <- chocolate %>%
  mutate(val_bin = case_when(
    cocoa_val < 68 ~ 'low',
    cocoa_val <= 72 ~ 'medium',
    cocoa_val > 72 ~ 'high'
  )) %>%
  mutate(val_bin = factor(val_bin,
                          levels=c('low', 'medium', 'high')))
```


```{r}
chocolate %>%
  select(review_date, rating, n_ingredients, cocoa_val, val_bin) %>%
  drop_na() %>%
ggplot(aes(cocoa_val, rating)) +
  geom_hex(binwidth = c(5,.25)) +
  facet_wrap(~n_ingredients)
```


```{r}
chocolate %>%
  mutate(region = case_when(
    company_location %in% c('Amsterdam',  'Czech Republic', 'Finland','Hungary',
                            'Iceland', 'Ireland', 'Lithuania', 'Netherlands', 
                            'Norway', 'Poland', 'Portugal',
                            'Scotland', 'Sweden', 'Wales') ~ 'Europe',
    company_location %in% c('Argentina', 'Bolivia', 'Brazil', 'Chile',
                            'Colombia', 'Costa Rica', 'Dominican Republic',
                            'El Salvador', 'Guatemala', 'Honduras',
                            'Nicaragua', 'Peru', 'Suriname') ~ 'South America',
    company_location %in% c('Fiji', 'Vanuatu') ~ 'Oceania',
    company_location %in% c('Sao Tome', 'Sao Tome & Principe') ~ 'Africa',
    company_location %in% c('Russia', 'Vietnam') ~ 'Asia',
    company_location %in% c('Grenada', 'St. Lucia', 'St.Vincent-Grenadines' ) ~ 'Caribbean',
    TRUE ~ NA_character_
  )) %>%
  filter(is.na(region)) %>%
  count(company_location)
```

```{r}
chocolate <- chocolate %>% 
  mutate(Vanilla = case_when(
    str_detect(ingredients, "V") ~ 1,
    TRUE ~ 0
  )) %>% 
  mutate(Salt = case_when(
    str_detect(ingredients, 'Sa') ~ 1,
    TRUE ~ 0
  )) %>%
  mutate(Sugar = case_when(
    is.na(ingredients) ~ NA_real_,
    str_detect(ingredients, 'S\\*') ~ 2,
    str_detect(ingredients, '[S,|S\\$]') ~ 1,
    TRUE ~ 0
  )) 
chocolate %>% count(Sugar, Vanilla, Salt)
```

```{r}
chocolate %>%
  select(rating, Sugar) %>%
  drop_na() %>% 
  ggplot(aes(rating, as.factor(Sugar), fill=as.factor(Sugar))) +
  ggridges::geom_density_ridges()
```


